Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-11 methods_paper/fabl-paper.tex
erased auxilliary files

\newpage
{
\color{red}
Here is the previous version of everything. Note that the numbering of the sections is off at the moment. 
}

%% Reworking introduction for Brian's paper.

Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. Many statistical record linkage methods and case studies are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test (or hypothesis test). Recent work in the statistical literature has extended the aforementioned work \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.

%While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease. Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here:

In this paper, we consider bipartite record linkage, which is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle_bayesian_2017}. Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

The Fellegi-Sunter method and its extensions are widely popular mainly due to its simplicity and computational scalability. Despite this, like any method, it has limitations, that are well known in the literature. Specifically, the assumption of no duplications within a database implies a maximum one-to-one constraint. That is, a record from one database can be linked with only one record in the other database. Modern methods of Fellegi-Sunter either ignore this restriction \citep{Winkler1988, belin_1995, larsen_2001}, include this as a post-processing step \citep{jaro1989}, or include it directly in the modeling framework \citep{sadinle_bayesian_2017}. It is important to note that \cite{fellegi_theory_1969} ignored this restriction as well. In practice, it seems that one should either directly include it in the modeling framework or correct for this in a post-processing step.

To our knowledge, \cite{sadinle_bayesian_2017} was the first to propose correcting the Fellegi and Sunter model for the one-to-one constraint. Specifically, the authors propose a Bayesian variant of the original model that satisfies two goals. First, it allows one to quantify uncertainty regarding match or non-match status. In addition, posterior distributions are readily available to help estimate point estimates of interest (and quantity such estimates). Second, under the Bayesian paradigm, the one-to-one matching constraint is easily embedded. The downside to this constraint is an added computational cost. The goal of our paper is proposing an extension of \cite{sadinle_bayesian_2017} that scales to large databases, but does not compromise much in terms of accuracy on the original case study considered by the author.

\textbf{Our Contribution} 
In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\texttt{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations and to reduce computational complexity.  Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching. Instead, we propose independent priors (breaking the one-to-one matching) in order to create computational speeds and gains. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices. 
To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
pronounced ``fable'')

The remainder of the paper is as follows. Section~\ref{sec:bipartite} reviews bipartite record linkage; section \ref{sec:fs} reviews the traditional approach of the Fellegi and Sunter framework before introducing our proposed work.

\section{Bipartite Record Linkage and Related Work}
As already mentioned, bipartite record linkage is the traditional set up studied in record linkage, where we assume there are two databases $\bm{X}_1$ and $\bm{X}_2$ such that there are duplications across the databases but not within each databases. We assume that each database has $n_1$ and $n_2$ records, and that $n_1 \leq n_2.$ Denote the number of entities simultaneously recorded by \emph{both databases} by $n_{12}$ such that $0 \leq n_{12} \leq n_2.$ Given this set up, as explained by \cite{steorts_bayesian_2016} and \cite{sadinle_bayesian_2017}, this can be viewed as bipartite record linkage. This is important as it allows one to view the matching matrix or linkage structure in the following way:
\begin{align}
\Delta_{ij} =
\begin{cases}
1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
0 \quad \text{otherwise}.\\
\end{cases}
\end{align}
Bipartite matching can also be viewed as a matching labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
Z_{j} =
\begin{cases}
i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

There are many approaches to bipartite record linkage, and we review some of the most common approaches below.

\paragraph{Fellegi and Sunter}
As already mentioned, the most common approach is the original approach of \cite{fellegi_theory_1969}. \cite{enamorado2019using} have extended this approach using computational speeds ups to allow for a much more scalable and practical approach, providing open source code for the community.

%These approaches have led to downstream tasks, such as regression approaches in the literature \citep{tang2020}.

\paragraph{Directly Modeling the Data} Instead of relying on comparison vectors, some authors will incorporate the data directly into their bipartite record linkage model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, gutman_bayesian_2013, shan2020bayesian}. These approaches have led to extensions regarding downstream tasks in the literature or joint models, such as regression approaches \citep{dalzell2018regression, steorts2018generalized, tancredi2020unified}.  There have been extensions to directly model the data in this literature where authors consider merging multiple databases (more than two) \citep{steorts_bayesian_2016, steorts_entity_2015, zanella_flexible_2016, marchant_distributed_2019, tancredi_2015_regression}. 
We do not review this literature given that it is not directly/fairly comparable with the case of bipartite record linkage.

\paragraph{Classification Approaches} 
One common way of approaching the record linkage task, especially in computer science and machine learning is using a classification approach. One uses a model (such as random forests) to classify records into matches/non-matches. When access to unique identifiers is available, one can create a training data (or reference data) set of record pairs (where the true match status is known).  Finally, one can train a classifier using the reference data set to predict the match status on any record pairs that were not included in the reference data set \citep{Bilenko2006, Christen2008, ventura2015seeing}.

\section{Review of the Fellegi Sunter Method}
In this section, we review notation, terminology, and the classical approach of Fellegi and Sunter.

Consider ordered record pairs $\bm{X}_1 \times \bm{X}_2,$ where the size of the databases are $n_1$ and $n_2,$ respectively. 
Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$ Denote the set of non-matches 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ An alternative way of merging two databases and removing 
duplications can be viewed as identifying the sets of matches/non-matches, or rather $\bm{M}, \bm{U}.$ In the literature, it is well known that record pairs that are estimated as matches are known as links; record pairs that are estimated as non-matches are known as non-links. \cite{fellegi_theory_1969} proposed using all-to-all pairwise comparison of records to estimate the matches/non-matches.

\subsection{Comparison Data}
In this section, we review the notion of \emph{comparison data}. Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar. A common way to encode this is using a comparison vector $\bm{\gamma}_{ij}$ which is computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote $F$ as the criteria for compare the records such that 
$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $F$ represents to a single comparison per feature that the databases share, such as gender. 
%
The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for nominal variables such as gender. For more complex measurements, such as textual or numeric feature information, we can take into account partial agreement patterns using common string metrics, such as the Edit, Jaro, or Jaro-Winkler distance functions. This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$ representing disagreement levels. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij} = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$ The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.

\subsection{Blocking}
In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, clusters, or ``blocks" to reduces this computational burden. Blocking can be deterministic, such as reducing the dimensionality of the comparison space on a highly reliable feature, such a gender. Blocking can be probabilistic, such as utilizing machine learning methods, known as hashing, which seek to use probabilistic functions (with low computational cost) to place similar records into the same bin. We propose hashing base methods for Bayesian bipartite record linkage, illustrating their ability to improve upon computational scalability. We emphasize that depending on the size of the problem, our proposed methodology can be used with or without hashing.

\subsection{Fellegi-Sunter Model}
In this section, we review the Fellegi-Sunter model. While we have introduced comparison data, $\bm{\gamma}_{ij}$, this is not sufficient to determine whether a record is a match/non-match given errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
\begin{align}
\label{eqn:wts}
w_{ij} = \log \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match/non-match. Assume $\bm{\Gamma}_{ij}$ is a random vector, whose distribution depends on $\Delta_{ij}.$ Then equation \ref{eqn:wts} is a representative value of $\bm{\Gamma}_{ij} \mid \Delta_{ij}.$ The ratio will be large if we favor the pair being a match. This is a problem in practice as transitive closures are violated. For example, suppose records $i$ and $j$ in $\bm{X}_1$ are extremely similar but are known to be non-matches. Suppose that both records $i$ and $j$ in $\bm{X}_1$ are very similar to record $k$ in $\bm{X}_2.$ By transitive closures, this implies that $i$ and $j$ are a match, which cannot be true by assumption.

\cite{jaro1989} proposed addressing this issue via an optimization approach as follows:
\begin{align}
\label{eqn:jaro}
&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
\end{align}

Equation \ref{eqn:jaro} ensures that each record of $\bm{X}_2$ only matches with at most one record in $\bm{X}_1$ (and the reverse condition). This is commonly known as the \emph{maximum-weight bipartite matching problem} or a \emph{linear sum assignment problem}. The output of this algorithm is a bipartite matching that maximizes the sum of the weights among matched pairs, where the pairs are not matched are assumed to be non-matches. \cite{jaro1989} provided no theoretical justification for using this approach, however, \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.

\subsection{Model Estimation}
In our review, we have assumed one knows $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0),$ however, these are typically not known in practice and must be estimated. Assume the comparison vectors are independent given the bipartite matching and assume that the match status of the record pairs are independent. It then follows that we can model 
the comparison data using mixture models in the following way:
\begin{align}
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match. Note that $\bm{m}$ and $\bm{u}$ are vectors of the match/non-match parameters. One common way to estimate the unknown parameters of the aforementioned model is using the EM algorithm.

One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures). One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases. On the other hand, one can turn to post-processing methods in practical situations. In this paper, we attempt to bridge the best of both approaches. Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}. We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step. In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}. Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.

\section{Proposed Methodology}
\label{sec:model}

In this section, we propose the fast beta linkage method (fabl) for bipartite record linkage, extending the work of \cite{sadinle_bayesian_2017}.  Specifically, we review the general approach of \cite{sadinle_bayesian_2017} and then provide our extensions for scalability.

%Specifically, the author used the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. In each iteration of his Gibbs sampler, the author considered each record \(j\in \bm{X}_2\), and then removed from consideration the records \(i\in \bm{X}_1\) that have already been matched. Then a potential link is sampled. While this leads to a sampler that is more accurate than the classical Fellegi-Sunter approach, accounting for such dependencies throughout the linkage process is computationally burdensome, and this method is only suitable for small databases. 
%As such, we propose a framework for efficient computation in large linkage tasks.

\subsection{Beta Linkage Model}
In this section, we review the Beta Linkage Model, which can be written as follows: 
%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings and $\mathcal{M}(\bm{m}), \mathcal{M}(\bm{u})$ are models for the comparison vectors of matches and non-matches.

\hypertarget{model-specification}{%
	\subsection{New Model Specification}
	\label{model-specification}}

In this section, we provide the specification for the models of $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u}),$ which follows from assuming that the features are conditionally independent of the level of agreement of another. The likelihood can be written as follows:

\begin{align}
\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}, 
\end{align}
where $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denotes the probability of a match have level $\ell$ of disagreement in feature $f$ and 
$u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represents the same probability for non-matches. 
In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, 
$\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.

Next, we allow for missing values or rather missing comparisons for corresponding record pairs. As commonly done in the record linkage literature, we assume that these occur missing at random [CITE: Little and Rubin (2002), Sadinle (2017), Steorts et al (2016)].

The likelihood becomes 
\begin{align}
\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}^{obs}) = 
\prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{a_{f\ell}(\bm{Z})}u_{fl}^{a_{f\ell}(\bm{Z})}\right], 
\end{align}
where 
\begin{align}
&a_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = \ell) I(Z_j = i) \\
&b_{f\ell}(\bm{Z}) = \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = \ell) I(Z_j \neq i),
\end{align}
where $I_{obs}$ denotes if a value is observed or not. Next, for a particular matching configuration $\bm{Z}$, $a_{f\ell}(\bm{Z})$
and $b_{f\ell}(\bm{Z})$ denote the number of matches and non-matches with observed disagreement level $\ell$ for comparison feature $f.$ Using this parameterization, we specify independent conjugate priors for all features $f$:
\begin{align}
&\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
&\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).
\end{align}

\hypertarget{fast-beta-prior}{%
	\subsection{Fast Beta Prior}
	\label{fast-beta-prior}}

In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$ While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.

\[Z_j | \lambda =
\begin{cases} 
	\frac{1}{n_1}\lambda  & z_j \leq n_1; \\
	1-\lambda &  z_j  = n_1 + j \\
\end{cases}\]

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \]

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data. One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).

There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}.

On the other hand, we propose independent priors for each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains. More importantly, since only the agreement pattern of \(Z_j\) is used for
calculations within the Gibbs sampler, and not the particular record
label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.  In doing so however, we thereby weaken the one-to-one requirement from
\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be
matched to at most one record in \(\bm{X}_1\), but allows for the possibility
that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).

\newpage

%\textcolor{red}{I think my notation is OK from earlier.}

%Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
%respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
%in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
%generality, label the files such that \(n_A \geq n_B\). We assume there
%are no duplicates within files, only across. For each record pair under
%consideration, we generate a comparison vector
%\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
%where \(F\) is the number of fields used in the linkage and each
%\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
%indicating the level agreement between the two records on a specified
%field.

>>>>

%\newpage
%\hypertarget{posterior-sampling}{%
%	\subsection{Gibbs Sampler}
%	\label{gibbs_sampling}}
%We work with the following factorization of the joint posterior distribution:
%\textcolor{red}{Brian: The Gibbs sampler has some minor typos. It should be written in terms of $\bm{Z}, \Phi, \lambda \mid \gamma_{obs}.$ Also, just make the rest of the notation match please:)}
%\begin{align*}
%p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
%&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)} \\
%&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
%&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
%\end{align*}
%This factorization leads to following Gibbs Sampler:
%\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
%The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
%standard multinomial-dirichlet mechanics. Thus we have
%\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
%\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
%where
%\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I(\gamma_{ij}^f = l) I(z_j = i)\)
%and
%\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + I(\gamma_{ij}^f = l) I(z_j \neq i)\).
%\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
%\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
%of successes (when \(z_j < n_A + 1\)) and failures (when
%\(z_j = n_A + 1\)), and therefore
%\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
%determined only by the number of duplicates
%$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
%$\mathbf{Z}$. Thus we have
%\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
%\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
%\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
%\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
%\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
%Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
%only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
%denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
%that as a function of \(Z_j\), the likelihood
%\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
%is a discrete distribution with probabilities proportional to
%\begin{align*}
%	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
%	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
%	&=
%	\begin{cases} 
%		w_{ij}  & z_j \leq n_1; \\
%		1 &  z_j  = n_1 + j \\
%	\end{cases}\\
%\end{align*}
%where
%\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
%The interested reader should note that these are precisely the
%likelihood ratios used in the Fellegi-Sunter model to classify matches
%and non-matches, and we therefore refer to \(w_{ij}\) as the
%\emph{Fellegi Sunter weights}.
%With the likelihood in this form, we can derive the full conditional
%\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]
%\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
%\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
%\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
%\begin{cases} 
%	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
%	1-\lambda &  z_j  = n_A + 1 \\
%\end{cases}\]
%In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
%integrate over the posterior of \(\lambda\) and rearrange terms to
%produce the final full conditional:
%\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
%\begin{cases} 
%	w_{ij}  & i \leq n_A; \\
%	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
%\end{cases}\]
%\hypertarget{bayes-estimate}{%
%	\subsection{Bayes Estimate}
%	\label{bayes-estimate}}
%We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
%parameter \(\mathbf{Z}\) by assigning different positive losses to
%different types of errors, and minimizing posterior expected loss. We
%adopt the loss function proposed in Sadinle 2017, in which
%\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
%option to leave the matching undetermined by the model. Specifically, we have
%\[L(\hat{Z_j}, Z_j)=\begin{cases} 
%	0  & \text{if } Z_j = \hat{Z_j}; \\
%	\theta_R,  & \text{if } \hat{Z_j} = R; \\
%	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
%	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
%	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
%\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
%on the linkage status, \(\theta_{10}\) is the loss from a false
%non-match, \(\theta_{01}\) is the loss from a false match, and
%\(\theta_{11}\) is the loss from the special case of a false match in
%which the record has a true match other than the one estimated by the
%model. 
%We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$. In this paper, we adopt losses
%\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
%\[\hat{Z}_j =\begin{cases} 
%	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
%	0,  & \text{otherwise} ; \\
%\end{cases}\]
%For a more in-depth explanation of this function and the induced Bayes
%estimate, see \citep{sadinle_bayesian_2017}.
%Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study. 
%To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that two matches. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
